---
title: "Meet the toolkit: Flint water crisis"
author: ""
date: ""
output: 
  html_document:
    toc: true
    number_sections: false
    toc_float: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = NA)
```

# Introduction

The data set consists of 271 homes sampled with three water 
lead contaminant values at designated time points. The lead content is in 
parts per billion (ppb). Additionally, some location data is given 
about each home.

To get started, read in the `flint.csv` file using function `read_csv()`.

```{r packages}
library(tidyverse)
```

```{r read-data}
flint <- read_csv("flint.csv")
```

Let's preview the data with function `glimpse()`

```{r}
glimpse(flint)
```

- **id**: sample id number
- **zip**: zip code in Flint as to the water sample's location
- **ward**: ward in Flint as to the water sample's location
- **draw**: water sample at one of three time points
- **lead**: lead content in parts per billion

# Analysis

## Summary data

#### Part I

Let's see how many samples were taken from each zip code.

```{r}
flint %>%               # data
  group_by(zip) %>%     # perform a grouping by zip code
  count()               # count occurrences
```

Which zip code had the most samples drawn?

#### Part II

Next, let's look at the mean and median lead contaminant values for each
zip code and draw combination. We have eight zip codes and samples taken at
three time points. How many combinations do we have?

```{r}
flint %>% 
  group_by(zip, draw) %>% 
  summarise(mean_pb = mean(lead))
```

```{r}
flint %>% 
  group_by(zip, draw) %>% 
  summarise(median_pb = median(lead))
```

How many rows are in each of two above data frames?

#### Part III

Modify the code below to compute the mean and median lead contaminant values
for zip code 48503 at the first draw. What should you put in for 
`draw == "-----"`? Don't forget to uncomment the second line of code.

```{r}
flint %>% 
  # filter(zip == 48503, draw == "-----") %>% 
  summarise(mean_pb = mean(lead),
            median_pb = median(lead))
```

## Visualizations

Let's make some plots, where we will focus on zip codes 48503, 48504, 48505,
48506, and 48507. We will restrict our attention to samples with lead values
less than 1,000 ppb.

```{r}
flint_focus <- flint %>% 
  filter(zip %in% 48503:48507, lead < 1000)
```

Below are side-by-side box plots for the three flushing times in each of the
five zip codes considered. Add `x` and `y` labels; add a title by inserting
`title = "title_name"` inside the `labs()` function.

```{r full-plot, fig.width=9}
ggplot(data = flint_focus, aes(x = factor(zip), y = lead)) +
  geom_boxplot(aes(fill = factor(draw))) +
  labs(x = "--------", y = "--------", fill = "Flushing time") +
  scale_fill_discrete(breaks = c("first", "second", "third"),
                      labels = c("0 (sec)", "45 (sec)", "120 (sec)")) +
  coord_flip() +
  theme_bw()
```

Add labels for `x`, `y`, a `title`, and `subtitle` to the code below to update
the corresponding plot.

```{r zoom-plot, fig.width=9}
ggplot(data = flint_focus, aes(x = factor(zip), y = lead)) +
  geom_boxplot(aes(fill = factor(draw))) + 
  labs(x = "--------", y = "--------", fill = "Flushing time",
       subtitle = "--------") +
  scale_fill_discrete(breaks = c("first", "second", "third"),
                      labels = c("0 (sec)", "45 (sec)", "120 (sec)")) +
  coord_flip(ylim = c(0, 50)) +
  theme_bw()
```

What is the difference between the two plots?

# References

1. Langkjaer-Bain, R. (2017), The murky tale of Flint's deceptive water data. 
   Significance, 14: 16-21.
